Finiteness in Cubical Type Theory

### 2020-01-20 paper.tex
more!

Finally, the other important property of constructive mathematics is that we can
actually \emph{compute} with proofs.
Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
even in the presence of univalence, giving computational content to programs
written in HoTT.
\subsection{Contributions}
In this work we will examine five notions of finiteness in Homotopy Type Theory,
the relationships between them, and their topos-theoretic characterisation.
We also briefly examine a definition for countable sets, comparing it to the
manifestly Bishop finite sets.
Our work is formalised in Cubical Agda, where we also develop a library for
proof search based on the finiteness predicates.

@@

All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in the
final section, where we implement a library for proof search, based on
omniscience and exhaustibility.
While proof search based on finiteness is not new, implementing it here does
give us an opportunity to demonstrate some actual \emph{computation} in a
univalent setting.
Many of the finiteness predicates are built using univalence, so for the proof
search functionality to work at all we require computational content in the
univalence axiom.
Furthermore, the extensionality afforded to us by HoTT means that we can have
\(\Pi\) types in the domain of the search, meaning we can synthesise full
functions from specifications alone.
\subsection{Related Work}
\subsection{Notation}
\todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
\todo[inline]{Paths, Univalence, isomorphisms}
\todo[inline]{HITs}
\todo[inline]{Surjections}
\todo[inline]{Containers}
\todo[inline]{Fin}
\todo[inline]{Biconditional, Isomorphism, Equivalence, Equals}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Kuratowski Finiteness}
Much work has already been done on Kuratowski finiteness in HoTT in
\cite{fruminFiniteSetsHomotopy2018}.
As a result, we will not needlessly repeat proofs, rather we will just give a
brief introduction to the topic and point out where our treatment differs.

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{aligned}
      \mathcal{K}(A) =&
      \; [] : \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \equiv y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \equiv x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \equiv \mathit{ys}) , p \equiv q ;
    \end{aligned}
  \end{equation}
\end{romdefinition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{romdefinition}[Membership of \(\mathcal{K}\)]
  Membership is defined by pattern-matching on \(\mathcal{K}\).
  The two point constructors are handled like so:
  \begin{equation}
    \begin{alignedat}{2}
      x \in&& \; []                      &= \bot \\
      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \equiv y \uplus x \in \mathit{ys} \rVert
    \end{alignedat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(\uplus\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{romdefinition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{romdefinition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{romdefinition}

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\subsection{Split Enumerable}
\begin{romdefinition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \equiv \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\todo[inline]{Put refs for each ``definition of''}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    & \text{Definition of } \mathcal{E}!
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    & \text{Definition of } \in
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of sp-surj}
    \\
    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}

@@

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    & \text{Definition of } \mathcal{B}
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    & \text{Definition of } \inunique
    \\
    & \equiv \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{\cite[theorem 4.4.5]{hottbook}}
    \\
    & \equiv \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \equiv \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \equiv \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \equiv \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}
\subsection{Manifest Enumerable}
\begin{romdefinition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{romdefinition}
Again, the only difference with this type and split enumerability is the
membership proof.
Here we have propositionally truncated it.

% We will use the circle as an example non-set type which is nonetheless
% manifestly enumerable.
% \begin{romtheorem}
%   The circle \(S^1\) is manifestly enumerable.
% \end{romtheorem}
% \begin{proof}
%   As the cover proof is a truncated proposition, we need only consider the point
%   constructors, making this poof the same as the proof of split enumerability on
%   \(\top\).
% \end{proof}
% \subsection{Surjections}
% This predicates relation to surjectivity is much the same as split
% enumerability's relation to \emph{split} surjectivity.
% \begin{romlemma}
%   A proof of manifest enumerability is equivalent to a surjection from a finite
%   prefix of the natural numbers.
% \end{romlemma}
\subsection{Cardinal}
\section{Relations Between Each Finiteness Definition}
\section{Topos}
\begin{itemize}
  \item Complete (pullbacks + terminal object)
\end{itemize}
\section{Countably Infinite Types}
\section{Search}

@@

% \begin{romtheorem}
%   Any cardinal finite type with a (decidable) total order is manifestly Bishop
%   finite.
% \end{romtheorem}
% \begin{proof}
%   This proof is quite involved, and will rely on several subsequent lemmas, so
%   we will give only its outline here.
%   \begin{itemize}
%     \item First, we will convert to manifest enumerability: knowing that the
%       underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
%       go from manifest enumerability to split enumerability
%       (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
%       finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
%     \item To convert to manifest enumerability, we need to provide a support
%       list: this cannot simply be the support list hidden under the truncation,
%       since that would violate the hiding promised by the truncation.
%       Instead, we sort the list (using insertion sort).
%       We must, therefore, prove that insertion sort is invariant under all
%       support lists in cardinal finiteness proofs.
%     \item We do this by first showing that all support lists in cardinal
%       finiteness proofs are permutations of each other, and then that insertion
%       sort is invariant under permutations.
%     \item Given our particular definition of permutations, cover proofs
%       transfers naturally between lists which are permutations of each other.
%   \end{itemize} 
% \end{proof}
% Now we will build up the toolkit we need to perform the above steps.
% First, permutations.
% \begin{romdefinition}[List Permutations]
%   We say that two lists are permutations of each other if there is an
%   isomorphism between membership proofs
%   \cite{danielssonBagEquivalenceProofRelevant2012}.
%   \begin{equation}
%     \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
%   \end{equation}
% \end{romdefinition}
% \begin{romlemma}
%   Insertion sort is invariant under permutations.
%   \begin{equation}
%     \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \equiv \text{sort}(\mathit{ys})
%   \end{equation}
% \end{romlemma}
% \begin{proof}
%   First, we prove two properties about insertion sort:
%   \begin{enumerate}
%     \item It returns a sorted list.
%     \item It a returns a list that is a permutation of its input.
%   \end{enumerate}
%   The second of these points allows us to show that \(\text{sort}(\mathit{xs})\)
%   is a permutation of \(\text{sort}(\mathit{ys})\).
%   \begin{equation}
%     \text{sort}(\mathit{xs}) \leftrightsquigarrow \mathit{xs} \leftrightsquigarrow \mathit{ys} \leftrightsquigarrow \text{sort}(\mathit{ys})
%   \end{equation}
%   Then, we show that any lists which are both sorted and permutations of each
%   other are equal.
%   Both of these conditions are true for the output of \(\text{sort}\).
% \end{proof}
% \section{Kuratowski Finiteness} \label{kuratowski}
% Finally we arrive at Kuratowski finiteness
% \cite{kuratowskiNotionEnsembleFini1920}.

>>>>

Finally, the other important property of constructive mathematics is that we can
actually \emph{compute} with proofs.
Cubical Type Theory \cite{cohenCubicalTypeTheory2016}, and its implementation in
Cubical Agda \cite{vezzosiCubicalAgdaDependently2019}, realise this property
even in the presence of univalence, giving computational content to programs
written in HoTT.
\subsection{Contributions}
In this work we will examine five notions of finiteness in Homotopy Type Theory,
the relationships between them, and their topos-theoretic characterisation.
We also briefly examine a predicate for countable sets, comparing it to the
finiteness predicates.
Our work is formalised in Cubical Agda, where we also develop a library for
proof search based on the finiteness predicates.

@@

All of our work is formalised in Cubical Agda
\cite{vezzosiCubicalAgdaDependently2019}.
We will make mention of the few occasions where the formalisation of some proof
is of interest, but the main place where we will discuss the code is in the
final section, where we implement a library for proof search, based on
omniscience and exhaustibility.
While proof search based on finiteness is not new, implementing it here does
give us an opportunity to demonstrate some actual \emph{computation} in a
univalent setting.
Many of the finiteness predicates are built using univalence, so for the proof
search functionality to work at all we require computational content in the
univalence axiom.
Furthermore, the extensionality afforded to us by HoTT means that we can have
\(\Pi\) types in the domain of the search, meaning we can synthesise full
functions from specifications alone.
\subsection{Related Work}
\subsection{Notation and Background}
We work in Cubical Type Theory \cite{cohenCubicalTypeTheory2016}.
For the various type formers we use the following notation:
\begin{description}
  \item[Type] We use \(\mathbf{Type}\) to denote the universe of (small) types.
    ``Type families'' are functions into type.
  \item[Universes]
    Implicitly, types will exist at some point in a universe hierarchy,
    beginning with \(\mathbf{Type}\) (\(\mathbf{Type} : \mathbf{Type}_1\),
    \(\mathbf{Type}_1 : \mathbf{Type}_2\), and so on).
    In our formalisation, every proof is done in the most universe polymorphic
    way possible: the proof that the cardinally finite sets form a
    \(\Pi\)-pretopos, for instance, is defined over any universe level.
  \item[0 , 1 , 2] We call the \(\mathbf{0}\), \(\mathbf{1}\), and
    \(\mathbf{2}\) types \(\bot\), \(\top\), and \(\mathbf{Bool}\) respectively.
    The single inhabitant of \(\top\) is tt, and the two inhabitants of
    \(\mathbf{Bool}\) are false and true.
  \item[Dependent Sum and Product] We use \(\Sigma\) and \(\Pi\) for the
    dependent sum and product, respectively.
    The two projections from \(\Sigma\) are called fst and snd.
    In the non-dependent case, \(\Sigma\) can be written as \(\times\), and
    \(\Pi\) as \(\rightarrow\).
  \item[Disjoint Union] Disjoint union can be defined in terms of \(\Sigma\):
    \begin{equation}
      A \uplus B \coloneqq \Sigma(x : \mathbf{Bool}) , \text{if } x \text{ then } A \text{ else } B
    \end{equation}
    However, we prefer to use it as an inductively defined type (though the two
    are equivalent).
    The two constructors are:
    \begin{equation}
      \begin{aligned}
        \text{inl} &: A \rightarrow A \uplus B \\
        \text{inr} &: B \rightarrow A \uplus B
      \end{aligned}
    \end{equation}
  \item[Equalities, equivalences, and paths] We use the symbol \(\coloneqq\)
    for definitions.
    \(\simeq\) will be used for equivalences, and \(\equiv\) for equalities.
    Of course, we know that \((A \simeq B) \simeq (A \equiv B)\) by univalence,
    so the distinction isn't terribly important in usage.
\end{description}

% \item[Isomorphisms and Biconditionals] A pair of functions \(f : A \rightarrow
%   B\) and \(g : B \rightarrow A\) will be described by the type \(A
%   \leftrightarrow B\).
%   If \(g\) is a quasi-inverse \cite[definition 2.4.6]{hottbook} of \(f\), we
%   have an isomorphism between \(A\) and \(B\), denoted \(A \iff B\).
%   Note that while we can derive an equivalence from an isomorphism, we do not
%   have \((A \iff B) \simeq (A \simeq B)\).

We will use the standard \(\Pi\) and \(\Sigma\) notation for dependent products
and sums, respectively.

\todo[inline]{Basic MLTT Notation: \(\Pi\), \(\Sigma\), etc}
\todo[inline]{Paths, Univalence, isomorphisms}
\todo[inline]{HITs}
\todo[inline]{Surjections}
\todo[inline]{Containers}
\todo[inline]{Fin}
\todo[inline]{Biconditional, Isomorphism, Equivalence, Equals}
\section{Finiteness Predicates}
In this section, we will define and briefly describe each of the five predicates
in Figure~\ref{finite-classification}.
The reason we explore predicates other than our focus (cardinal finiteness) is
that we can often prove things like closure much more readily on the simpler
predicates.
The relations (which we will prove in the next section) then allow us to
transfer those proofs onto Kuratowski finiteness.
\subsection{Split Enumerable}
\begin{romdefinition}[Split Enumerable Set]
  \begin{equation}
    \mathcal{E}!(A) = \Sigma {(\mathit{xs} : \textbf{List}(A))} , \Pi {(x : A)} , x \in \mathit{xs}
  \end{equation}
  We call the first component of this pair the ``support'' list, and the second
  component the ``cover'' proof.
  An equivalent version of this predicate was called \verb+Listable+ in
  \cite{firsovDependentlyTypedProgramming2015}.
\end{romdefinition}

We tend to prefer list-based definitions of finiteness, rather than ones based
on bijections or surjections.
This is purely a matter of perspective, however: the definition above is
precisely equivalent to a split surjection from a finite prefix of the natural
numbers.
\begin{romlemma} \label{split-enum-is-split-surj}
  \begin{equation}
    \mathcal{E}!(A) \simeq \Sigma (n : \mathbb{N}) , \left( \mathbf{Fin}(n) \twoheadrightarrow ! \; A \right)
  \end{equation}
\end{romlemma}
\todo[inline]{Put refs for each ``definition of''}
\begin{proof}
  \begin{align*}
    \mathcal{E}!(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \in \mathit{xs}
    & \text{Definition of } \mathcal{E}!
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{fib}_{\text{snd}(\mathit{xs})}(x)
    & \text{Definition of } \in
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of sp-surj}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{sp-surj}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{sp-surj}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \twoheadrightarrow ! \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}

@@

Where split enumerability was the enumeration form of a split surjection from
\(\mathbf{Fin}\), manifest Bishop finiteness is the enumeration form of an
\emph{equivalence} with \(\mathbf{Fin}\).
\begin{romlemma} \label{bishop-equiv}
  \begin{equation}
    \mathcal{B}(A) \simeq \Sigma {(n : \mathbb{N})} , \left( \mathbf{Fin}(n) \simeq A \right)
  \end{equation}
\end{romlemma}
\begin{proof}
  \begin{align*}
    \mathcal{B}(A)
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    & \text{Definition of } \mathcal{B}
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr} \left(\text{fib}_{\text{snd}(\mathit{xs})}(x)\right)
    & \text{Definition of } \inunique
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{\cite[theorem 4.4.5]{hottbook}}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \mathbf{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \mathbf{List}
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \mathbf{Fin}(n)) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    & \text{Definition of } \llbracket \wc \rrbracket
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \mathbf{Fin}(n) \rightarrow A) , \text{isEquiv}(f)
    & \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \mathbf{Fin}(n) \simeq \; A )
    & \text{Definition of } \twoheadrightarrow!
  \end{align*}
  \qed
\end{proof}
\subsection{Cardinal}
\begin{itemize}
\item Definition
\item Cardinality
\end{itemize}
\subsection{Manifest Enumerable}
\begin{romdefinition}[Manifest Enumerability]
  \begin{equation}
    \mathcal{E}(A) = \Sigma {(\mathit{xs} : \mathbf{List}(A))} , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
  \end{equation}
\end{romdefinition}
Again, the only difference with this type and split enumerability is the
membership proof: here we have propositionally truncated it.
This has two effects.
First, it means that this proof represents a true surjection (rather than a
split surjection) from \(\mathbf{Fin}\).
\begin{romlemma}
  \begin{equation}
    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\mathbf{Fin}(n) \twoheadrightarrow A)
  \end{equation}
\end{romlemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}

Secondly, it means the predicate does not imply decidable equality.
More significantly, it allows the predicate to be defined over non-set types,
like the circle.
\begin{romlemma}
  The circle \(S^1\) is manifestly enumerable.
\end{romlemma}
\begin{proof}
  \todo[inline]{Proof}
\end{proof}

\subsection{Kuratowski Finiteness}
Much work has already been done on Kuratowski finiteness in HoTT in
\cite{fruminFiniteSetsHomotopy2018}.
As a result, we will not needlessly repeat proofs, rather we will just give a
brief introduction to the topic and point out where our treatment differs.

The first thing we must define is a representation of subsets.
\begin{romdefinition}[Kuratowski Finite Subset]
  \(\mathcal{K}(A)\) is the type of Kuratowski-finite subsets of \(A\).
  \begin{equation}
    \begin{aligned}
      \mathcal{K}(A) =&
      \; [] : \mathcal{K}(A) ; \\
      \vert& \; \wc \dblcolon \wc : A \rightarrow \mathcal{K}(A) \rightarrow \mathcal{K}(A) ; \\
      \vert& \; \text{com} : \Pi (x, y: A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon y \dblcolon \mathit{xs} \simeq y \dblcolon x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{dup} : \Pi (x : A) , \Pi (\mathit{xs} : \mathcal{K}(A)) , x \dblcolon x \dblcolon \mathit{xs} \simeq x \dblcolon \mathit{xs} ; \\
      \vert& \; \text{trunc} : \Pi (\mathit{xs}, \mathit{ys}: \mathcal{K}(A)) , \Pi (p, q : \mathit{xs} \simeq \mathit{ys}) , p \simeq q ;
    \end{aligned}
  \end{equation}
\end{romdefinition}
The Kuratowski finite subset is a free join semilattice (or, equivalently, a
free commutative idempotent monoid).
More prosaically, \(\mathcal{K}\) is the abstract data type for finite sets, as
defined in the Boom hierarchy \cite{boomFurtherThoughtsAbstracto1981,
  bunkenburgBoomHierarchy1994}.
However, rather than just being a specification, \(\mathcal{K}\) is fully usable
as a data type in its own right, thanks to HITs.

Next, we need a way to say that an entire type is Kuratowski finite.
For that, we will need to define membership of \(\mathcal{K}\).
\begin{romdefinition}[Membership of \(\mathcal{K}\)]
  Membership is defined by pattern-matching on \(\mathcal{K}\).
  The two point constructors are handled like so:
  \begin{equation}
    \begin{alignedat}{2}
      x \in&& \; []                      &= \bot \\
      x \in&& \; y \dblcolon \mathit{ys} &= \lVert x \simeq y \uplus x \in \mathit{ys} \rVert
    \end{alignedat}
  \end{equation}
  The \(\text{com}\) and \(\text{dup}\) constructors are handled by proving that
  the truncated form of \(\uplus\) is itself commutative and idempotent.
  The type of propositions is itself a set, satisfying the \(\text{trunc}\)
  constructor.
\end{romdefinition}
Finally, we have enough background to define Kuratowski finiteness.
\begin{romdefinition}[Kuratowski Finiteness]
  \begin{equation}
    \mathcal{K}^{f}(A) = \Sigma {(\mathit{xs} : \mathcal{K}(A))} , \Pi (x : A) , x \in \mathit{xs}
  \end{equation}
\end{romdefinition}

We also have the following two lemmas, proven in both
\cite{fruminFiniteSetsHomotopy2018} and our formalisation.
\begin{romlemma}
  \(\mathcal{K}^f\) is a mere proposition.
\end{romlemma}
\begin{romlemma}
  This circle \(S^1\) is Kuratowski finite.
\end{romlemma}
The second of these in particular tells us that the ``Kuratowski-finite types''
are not necessarily sets; it also tells us that we cannot derive decidable
equality from a proof of Kuratowski finiteness.
\section{Relations Between Each Finiteness Definition}
\begin{itemize}
  \item SE -> MB (and back)
  \item MB <-> ME
  \item MB <-> C
  \item C <-> K
  \item ME -> K
\end{itemize}
\section{Topos}
\begin{itemize}
  \item \(\bot\)
  \item \(\top\)
  \item Bool
  \item \(\Sigma\)
  \item \(\Pi\)
  \item Pre-categories to categories
  \item \(\Pi\)-pretopos
\end{itemize}
\section{Countably Infinite Types}
\begin{itemize}
  \item Definition
  \item \(\Sigma\)
  \item Kleene star
  \item Truncated
  \item Topos
\end{itemize}
\section{Search}
\begin{itemize}
  \item Omniscience
  \item Search
  \item Curry / uncurry
  \item Synth function
\end{itemize}

@@

% \begin{romtheorem}
%   Any cardinal finite type with a (decidable) total order is manifestly Bishop
%   finite.
% \end{romtheorem}
% \begin{proof}
%   This proof is quite involved, and will rely on several subsequent lemmas, so
%   we will give only its outline here.
%   \begin{itemize}
%     \item First, we will convert to manifest enumerability: knowing that the
%       underlying type is discrete (theorem~\ref{cardinal-finite-discrete}) we can
%       go from manifest enumerability to split enumerability
%       (lemma~\ref{manifest-enum-to-split}), and subsequently to manifest Bishop
%       finiteness (lemma~\ref{split-enum-to-manifest-bishop}).
%     \item To convert to manifest enumerability, we need to provide a support
%       list: this cannot simply be the support list hidden under the truncation,
%       since that would violate the hiding promised by the truncation.
%       Instead, we sort the list (using insertion sort).
%       We must, therefore, prove that insertion sort is invariant under all
%       support lists in cardinal finiteness proofs.
%     \item We do this by first showing that all support lists in cardinal
%       finiteness proofs are permutations of each other, and then that insertion
%       sort is invariant under permutations.
%     \item Given our particular definition of permutations, cover proofs
%       transfers naturally between lists which are permutations of each other.
%   \end{itemize} 
% \end{proof}
% Now we will build up the toolkit we need to perform the above steps.
% First, permutations.
% \begin{romdefinition}[List Permutations]
%   We say that two lists are permutations of each other if there is an
%   isomorphism between membership proofs
%   \cite{danielssonBagEquivalenceProofRelevant2012}.
%   \begin{equation}
%     \mathit{xs} \leftrightsquigarrow \mathit{ys} = \Pi {(x : A)} , x \in \mathit{xs} \iff x \in \mathit{ys}
%   \end{equation}
% \end{romdefinition}
% \begin{romlemma}
%   Insertion sort is invariant under permutations.
%   \begin{equation}
%     \mathit{xs} \leftrightsquigarrow \mathit{ys} \implies \text{sort}(\mathit{xs}) \simeq \text{sort}(\mathit{ys})
%   \end{equation}
% \end{romlemma}
% \begin{proof}
%   First, we prove two properties about insertion sort:
%   \begin{enumerate}
%     \item It returns a sorted list.
%     \item It a returns a list that is a permutation of its input.
%   \end{enumerate}
%   The second of these points allows us to show that \(\text{sort}(\mathit{xs})\)
%   is a permutation of \(\text{sort}(\mathit{ys})\).
%   \begin{equation}
%     \text{sort}(\mathit{xs}) \leftrightsquigarrow \mathit{xs} \leftrightsquigarrow \mathit{ys} \leftrightsquigarrow \text{sort}(\mathit{ys})
%   \end{equation}
%   Then, we show that any lists which are both sorted and permutations of each
%   other are equal.
%   Both of these conditions are true for the output of \(\text{sort}\).
% \end{proof}
% \section{Kuratowski Finiteness} \label{kuratowski}
% Finally we arrive at Kuratowski finiteness
% \cite{kuratowskiNotionEnsembleFini1920}.
